---
pdf_document:
  toc: yes
author: "Andrew D. Johnston"
date: "08/31/2020"
output:
  html_document: null
  css: styles.css
  pdf_document: default
title: "Is Cycle Threshold A Worthwhile Clinical Metric for SARS-CoV2?"
editor_options:
  chunk_output_type: console
toc: yes
toc_depth: 4
word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
#getwd("/Users/andrew/Dropbox/COVID-19/research/")
```

Goals of Analysis

1) Calculate the Ct value difference between Expired and Survivors for each Machine - Cepheid, Hologic, and Abbott
2) Calculate the Youden index for each Machine - Cepheid, Hologic, and Abbott
3) Examine the variability of samples taken within 48 hours of each other.
4) Observe how many specimen pairs cross the Youden Index for their respective machine

```{r packages, echo=TRUE, warning=FALSE, message = FALSE}
# load in libraries
library(data.table)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(ggpubr)
library(RColorBrewer)
library(MASS)
library(AER) 
library(scales)
library(Hmisc)
library("flexplot")
library(overlapping)
library(lattice)
library(lubridate)
library(ROCit)
library(pROC)
library(plotROC)
library(OptimalCutpoints)
library(cutpointr)
library(pastecs)

# set options
options(scipen=999, stringsAsFactors = FALSE)
```

Loading in the mortatlity data for each COVID specimen

```{r mort-data-load, echo=TRUE, warning=FALSE, message = FALSE}
mort_dat <- fread("Data_used/COVID_POS_SURVIVAL_short.csv",  header = T)
mort_dat <- as.data.frame(mort_dat)
dim(mort_dat)
mort_dat$Alive <- "Survivor"
mort_dat$Alive[mort_dat$`Patient Age`=="Deceased"] <- "Expired"
mort_dat$Alive <- factor(mort_dat$Alive, levels = c("Survivor", "Expired"))
```

1a) Obtaining the CT differences between expired and survivor for Cephied Machine

```{r cephied-ct, echo=TRUE, warning=FALSE, message = FALSE}
cephied_dat <- fread("Data_used/sarsCoV2-cepheid_ct.csv", header = T)
cephied_dat <- as.data.frame(cephied_dat)
cephied_dat <- cephied_dat[1:11490,] # R read in all the empty rows
cephied_dat <- cephied_dat[,-c(4,6,8)] # R read in all the empty rows
colnames(cephied_dat) <- c("MRN", "Spec_ID", "Result", "Ct_E", "Ct_n", "Ct_Control","start_time")
head(cephied_dat)
dim(cephied_dat)

# any with same spec ID - QC
any(duplicated(cephied_dat$Spec_ID))

# clear all negative and error and  
filter_cepheid <-c("NEGATIVE","ERROR","INVALID", "NO RESULT")
idx_cephied_filter <- grep(paste0(filter_cepheid, collapse = "|"), x = cephied_dat$Result, value = FALSE)
length(idx_cephied_filter) #9589
cephied_dat <- cephied_dat[-idx_cephied_filter,]
table(cephied_dat$Result)
# there are 11 samples in which the control probe did not result but they were deemed positive anyways, so I'll keep them
cephied_dat[which(cephied_dat$Ct_Control==0),] 
# head(cephied_dat[cephied_dat$Result=="SARS-CoV-2 PRESUMPTIVE POS",],50)
# head(cephied_dat[which(cephied_dat$Ct_E==0),])
# the presumptive positives were when E probe was positive but N probe was undetected
# removing Ct_N value of 0 
cephied_dat <- cephied_dat[-which(cephied_dat$Ct_n==0),]
dim(cephied_dat) # 1828

cephied_dat_mort <- merge(x=cephied_dat, y=mort_dat, by.x = "Spec_ID", by.y = "Specimen ID")
dim(cephied_dat_mort) # lost two in the merge; 1826

wilcox.test(cephied_dat_mort$Ct_n[cephied_dat_mort$Alive=="Survivor"],cephied_dat_mort$Ct_n[cephied_dat_mort$Alive=="Expired"], conf.int=TRUE)

# 95 percent confidence interval:
# 4.300053 6.299982
# difference in location 
#              5.299954 

median(cephied_dat_mort$Ct_n[cephied_dat_mort$Alive=="Survivor"])-median(cephied_dat_mort$Ct_n[cephied_dat_mort$Alive=="Expired"])
mean(cephied_dat_mort$Ct_n[cephied_dat_mort$Alive=="Survivor"])-mean(cephied_dat_mort$Ct_n[cephied_dat_mort$Alive=="Expired"])


colorBlindBlack8  <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
pie(rep(1, 8), col = colorBlindBlack8)

dark_col <- brewer.pal(6,"Dark2")
fig_ct_method_cephied <- ggplot(cephied_dat_mort, aes(x=Alive, y=Ct_n, fill=Alive)) + 
geom_dotplot(binaxis='y', stackdir='center', stackratio=.5, dotsize=.25, col=NA, position = position_jitterd(width = .05, height = 1, quad.points = 1, seed = 7)) +
  theme_minimal() + 
  scale_fill_manual(values=colorBlindBlack8[c(4,6)]) +
  stat_summary(data = cephied_dat_mort, mapping = aes(x=as.numeric(as.factor((Alive))), y=Ct_n), fun.y= "median", geom="point", color=colorBlindBlack8[7], inherit.aes = FALSE, shape = 95, size=20) +
  xlab("")+
  ylab("Ct Value") +
  ylim(c(0,50))
fig_ct_method_cephied
#ggsave(plot = fig_ct_method_cephied, filename = "Fig1_cephied.pdf", width=4.5, height=3, useDingbats=FALSE)

# plot AUC overlap between survivor and expired
AUC_overlap_cephied <- list(Survivor=cephied_dat_mort$Ct_n[cephied_dat_mort$Alive=="Survivor"], Expired=cephied_dat_mort$Ct_n[cephied_dat_mort$Alive=="Expired"])
AUC_overlap_plot_cephied <- overlap(AUC_overlap_cephied, plot=TRUE, theme = theme_minimal)

fig_ct_overlap_cephied <- ggplot(AUC_overlap_plot_cephied$DD, aes(x,y1)) +
  geom_ribbon(aes(ymin=0,ymax=y2),alpha=.3,fill=colorBlindBlack8[c(6)])+
  geom_ribbon(aes(ymin=0,ymax=y1),alpha=.3,fill=colorBlindBlack8[c(4)])+
  xlab("Cycle Threshold")+
  xlim(c(0,50)) +
  ylab("") +
  theme_minimal()+
  ylim(c(0,0.08))+
  ggtitle(paste0("Overlap = ",round(AUC_overlap_plot_cephied$OV*100,2),"%"))
fig_ct_overlap_cephied
#ggsave(plot = fig_ct_overlap_cephied, filename = "Fig1_overlap_cephied.pdf", width=2.5, height=2.5, useDingbats=FALSE)
```

2a) Obtaining the Youden Index for the Cephied machine

```{r cephied-roc, echo=TRUE, warning=FALSE, message = FALSE}

rocfit_cepheid <- pROC::roc(cephied_dat_mort$Alive, cephied_dat_mort$Ct_n)
pROC::auc(rocfit_cepheid)
youden_coord_cephied<- pROC::coords(rocfit_cepheid, x = "b")
summary(cephied_dat_mort$Ct_n)
rocplot_cepheid <- ggplot(cephied_dat_mort, aes(m = Ct_n, d = Alive)) + 
  geom_roc(n.cuts=0)
rocplot_cepheid <- rocplot_cepheid + style_roc(theme = theme_minimal, xlab = "1 - Specificity", ylab = "Sensitivity") +
  annotate("text", x = .75, y = .25, 
           label = paste("AUC =", round(calc_auc(rocplot_cepheid)$AUC, 2)))  +
  geom_point(aes(x=1-youden_coord_cephied$specificity, y=youden_coord_cephied$sensitivity), colour="red")+
  geom_abline(slope=1, intercept=0, lty="dashed")
  #ggtitle("Cepheid ROC")
rocplot_cepheid
#ggsave(plot = rocplot_cepheid, filename = "Fig1_roc_cephied.pdf", width=2.5, height=2.5, useDingbats=FALSE)
```

1b) Obtaining the CT differences between expired and survivor for Hologic Machine

```{r hologic-ct, echo=TRUE, warning=FALSE, message = FALSE}
hologic_dat <- fread("Copy of pcrbase.csv", header = T)
hologic_dat <- as.data.frame(hologic_dat)
dim(hologic_dat) #16732

#sanity check- duplicated specimen IDs?
any(duplicated(hologic_dat$Spec_name))
sum(duplicated(hologic_dat$Spec_name)) #412

#removing duplicates
hologic_dat <- hologic_dat[!duplicated(hologic_dat$Spec_name),]

# sort based on MRN then time collected
hologic_dat <- hologic_dat[order(hologic_dat$mrn),]

# making MRN numeric 
hologic_dat$mrn <- as.numeric(hologic_dat$mrn)

# remove the hologi as well as NA MRN patients
hologic_dat <- hologic_dat[-c(1:2),]
sum(is.na(hologic_dat$mrn)) #130
hologic_dat <- hologic_dat[!is.na(hologic_dat$mrn),]

#removing negative tests
table(hologic_dat$Result)
hologic_dat <- hologic_dat[hologic_dat$Result=="POS",]

#QC 
table(hologic_dat$IC_valid)
summary(hologic_dat$Control_Ct) 
dim(hologic_dat)
hologic_dat_mort <- merge(x=hologic_dat, y=mort_dat, by.x = "Spec_name", by.y = "Specimen ID")
dim(hologic_dat_mort)

wilcox.test(hologic_dat_mort$Ct[hologic_dat_mort$Alive=="Survivor"],hologic_dat_mort$Ct[hologic_dat_mort$Alive=="Expired"], conf.int=TRUE)

# 95 percent confidence interval:
# 3.499936 5.000038
# sample estimates:
# difference in location 
#              4.299973

fig_ct_method_hologic <- ggplot(hologic_dat_mort, aes(x=Alive, y=Ct, fill=Alive)) + 
geom_dotplot(binaxis='y', stackdir='center', stackratio=.5, dotsize=.17, col=NA, position = position_jitterd(width = .05, height = 1, quad.points = 1, seed = 7)) +
  theme_minimal() + 
  scale_fill_manual(values=colorBlindBlack8[c(4,6)]) +
  stat_summary(data = hologic_dat_mort, mapping = aes(x=as.numeric(as.factor((Alive))), y=Ct), fun.y= "median", geom="point", color=colorBlindBlack8[7], inherit.aes = FALSE, shape = 95, size=20) +
  xlab("")+
  ylab("Ct Value") +
  ylim(c(0,50))
fig_ct_method_hologic
#ggsave(plot = fig_ct_method_hologic, filename = "Fig1_hologic.pdf", width=4.5, height=3, useDingbats=FALSE)

# plot AUC overlap between survivor and expired
AUC_overlap_hologic <- list(Survivor=hologic_dat_mort$Ct[hologic_dat_mort$Alive=="Survivor"], Expired=hologic_dat_mort$Ct[hologic_dat_mort$Alive=="Expired"])
AUC_overlap_plot_hologic <- overlap(AUC_overlap_hologic, plot=TRUE)

fig_ct_overlap_hologic <- ggplot(AUC_overlap_plot_hologic$DD, aes(x,y1)) +
  geom_ribbon(aes(ymin=0,ymax=y2),alpha=.3,fill=colorBlindBlack8[c(6)])+
  geom_ribbon(aes(ymin=0,ymax=y1),alpha=.3,fill=colorBlindBlack8[c(4)])+
  xlab("Cycle Threshold")+
  xlim(c(0,50)) +
  ylab("") +
  ylim(c(0,0.08))+
  theme_minimal()+
  ggtitle(paste0("Overlap = ",round(AUC_overlap_plot_hologic$OV*100,2),"%"))
fig_ct_overlap_hologic
#ggsave(plot = fig_ct_overlap_hologic, filename = "Fig1_overlap_hologic.pdf", width=2.5, height=2.5, useDingbats=FALSE)

```

2b) Obtaining the Youden Index for the Hologic machine


```{r hologic-roc, echo=TRUE, warning=FALSE, message = FALSE}

rocfit_hologic <- pROC::roc(as.character(hologic_dat_mort$Alive), hologic_dat_mort$Ct)
pROC::auc(rocfit_hologic)
youden_coord_hologic<- pROC::coords(rocfit_hologic, x = "b")
youden_coord_hologic
rocplot_hologic<- ggplot(hologic_dat_mort, aes(m = Ct, d = Alive)) + 
  geom_roc(n.cuts=0)
rocplot_hologic <- rocplot_hologic + style_roc(theme = theme_minimal, xlab = "1 - Specificity", ylab = "Sensitivity") +
  annotate("text", x = .75, y = .25, 
           label = paste("AUC =", round(calc_auc(rocplot_hologic)$AUC, 2)))  +
  geom_point(aes(x=1-youden_coord_hologic$specificity, y=youden_coord_hologic$sensitivity), colour="red")+
  geom_abline(slope=1, intercept=0, lty="dashed")
  #ggtitle("Hologic ROC")
rocplot_hologic
#ggsave(plot = rocplot_hologic, filename = "Fig1_roc_hologic.pdf", width=2.5, height=2.5, useDingbats=FALSE)
```

1c) Obtaining the CT differences between expired and survivor for Hologic Machine

```{r abbott-ct, echo=TRUE, warning=FALSE, message = FALSE}
abbott_dat <- fread("Ct_study_race.csv", header = T)
abbott_dat <- as.data.frame(abbott_dat)
dim(abbott_dat)

# sanity checks 
## are there duplicates? No
any(duplicated(abbott_dat$MR)) # FALSE
any(duplicated(abbott_dat$`Specimen ID`)) # FALSE

# how many samples don't have a CT value
sum(is.na(abbott_dat$Ct)) # 2
#  How many samples have ct of 0? 
sum(abbott_dat$Ct==0, na.rm = TRUE) # 1

# I'm removing these four  samples
abbott_dat <- abbott_dat[!(is.na(abbott_dat$Ct)) & !(abbott_dat$Ct==0),]
dim(abbott_dat)

# remove unecessary columns (Patient Name, Component, Column 3, method)
colnames(abbott_dat)
abbott_dat <- abbott_dat[,-c(3:4,7,21:27)]

# only using Abbott data points
abbott_dat <-subset(abbott_dat, `Method Abbr`=="m2000")
dim(abbott_dat)

# sort based on MRN then time collected
abbott_dat <- abbott_dat[order(abbott_dat$MR),]

# making MRN numeric 
abbott_dat$MR <- as.numeric(abbott_dat$MR)
sum(is.na(abbott_dat$mrn))

table(abbott_dat$Value)
summary(abbott_dat$Ct)

#combining with mortaility data
dim(abbott_dat)
abbott_dat_mort <- merge(x=abbott_dat, y=mort_dat, by.x = "Specimen ID", by.y = "Specimen ID")
dim(abbott_dat_mort) # all were kept :-)
colnames(abbott_dat_mort)[22] <- "Alive"

wilcox.test(abbott_dat_mort$Ct[abbott_dat_mort$Alive=="Survivor"],abbott_dat_mort$Ct[abbott_dat_mort$Alive=="Expired"], conf.int=TRUE)

# 95 percent confidence interval:
# 1.059992 2.669944
# sample estimates:
# difference in location 
#              1.859981 

fig_ct_method_abbott <- ggplot(abbott_dat_mort, aes(x=Alive, y=Ct, fill=Alive)) + 
geom_dotplot(binaxis='y', stackdir='center', stackratio=.5, dotsize=.25, col=NA, position = position_jitterd(width = .05, height = 1, quad.points = 1, seed = 7)) +
  theme_minimal() + 
  scale_fill_manual(values=colorBlindBlack8[c(4,6)]) +
  stat_summary(data = abbott_dat_mort, mapping = aes(x=as.numeric(as.factor((Alive))), y=Ct), fun.y= "median", geom="point", color=colorBlindBlack8[7], inherit.aes = FALSE, shape = 95, size=20) +
  xlab("")+
  ylab("Ct Value") +
  ylim(c(0,50))
fig_ct_method_abbott
#ggsave(plot = fig_ct_method_abbott, filename = "Fig1_abbott.pdf", width=4.5, height=3, useDingbats=FALSE)


# plot AUC overlap between survivor and expired
AUC_overlap_abbott <- list(Survivor=abbott_dat_mort$Ct[abbott_dat_mort$Alive=="Survivor"], Expired=abbott_dat_mort$Ct[abbott_dat_mort$Alive=="Expired"])
AUC_overlap_plot_abbott <- overlap(AUC_overlap_abbott, plot=TRUE, theme = theme_minimal)

fig_ct_overlap_abbott <- ggplot(AUC_overlap_plot_abbott$DD, aes(x,y1)) +
  geom_ribbon(aes(ymin=0,ymax=y2),alpha=.3,fill=colorBlindBlack8[c(6)])+
  geom_ribbon(aes(ymin=0,ymax=y1),alpha=.3,fill=colorBlindBlack8[c(4)])+
  xlab("Cycle Threshold")+
  xlim(c(0,50)) +
  ylab("") +
  ylim(c(0,0.08))+
  theme_minimal()+
  ggtitle(paste0("Overlap = ",round(AUC_overlap_plot_abbott$OV*100,2),"%"))
fig_ct_overlap_abbott
#ggsave(plot = fig_ct_overlap_abbott, filename = "Fig1_overlap_abbott.pdf", width=2.5, height=2.5, useDingbats=FALSE)
```

2c) Obtaining the Youden Index for the Abbott machine

```{r abbott-roc, echo=TRUE, warning=FALSE, message = FALSE}

rocfit_abbott <- pROC::roc(as.character(abbott_dat_mort$Alive), abbott_dat_mort$Ct)
pROC::auc(rocfit_abbott)
youden_coord_abbott<- pROC::coords(rocfit_abbott, x = "b")
youden_coord_abbott
rocplot_abbott<- ggplot(abbott_dat_mort, aes(m = Ct, d = Alive)) + 
  geom_roc(n.cuts=0)
rocplot_abbott <- rocplot_abbott + style_roc(theme = theme_minimal, xlab = "1 - Specificity", ylab = "Sensitivity") +
  annotate("text", x = .75, y = .25, 
           label = paste("AUC =", round(calc_auc(rocplot_abbott)$AUC, 2)))  +
  geom_point(aes(x=1-youden_coord_abbott$specificity, y=youden_coord_abbott$sensitivity), colour="red")+
  geom_abline(slope=1, intercept=0, lty="dashed")
  #ggtitle("abbott ROC")
rocplot_abbott
#ggsave(plot = rocplot_abbott, filename = "Fig1_roc_abbott.pdf", width=2.5, height=2.5, useDingbats=FALSE)
```

3) Examine the variability of samples taken within 48 hours of each other. reading and curating data from pairs of Ct value from Abbott and Cephied 

```{r pairs-data, echo=TRUE, warning=FALSE, message = FALSE}
pairs_dat <- fread("Data_used/within_48.csv", header = T)
pairs_dat <- as.data.frame(pairs_dat)
dim(pairs_dat) #93 9

# how many different machines
table(pairs_dat$Method)

#sanity check- duplicated specimen IDs?
any(duplicated(pairs_dat$Spec_name)) # FALSE

#parsing the date and time information
pairs_dat$collect_parse <- parse_date_time(pairs_dat$Collected, orders="m/d/y H:M p")

# sort based on MRN then time collected
pairs_dat <- pairs_dat[order(pairs_dat$mrn, pairs_dat$collect_parse),]

# how many patients have more than 2 samples? # two have 3, one has 5
table(table(pairs_dat$mrn))

# split into list
split_pairs <- split(pairs_dat, pairs_dat$mrn)

pairs_df <- data.frame() 
j<-1
for (i in 1:length(split_pairs)){
  if(nrow(split_pairs[[i]])<3){
    pairs_df[j,1] <- split_pairs[[i]]$mrn[1]
    pairs_df[j,2] <- diff(split_pairs[[i]]$Ct)
    pairs_df[j,3] <- as.numeric(difftime(split_pairs[[i]]$collect_parse[2],split_pairs[[i]]$collect_parse[1], units="hours"))
    pairs_df[j,4] <- split_pairs[[i]]$Age[1]
    pairs_df[j,5] <- split_pairs[[i]]$Sex[1]
    pairs_df[j,6] <- split_pairs[[i]]$Method[1]
    pairs_df[j,7] <- (split_pairs[[i]]$Method[1]==split_pairs[[i]]$Method[2])
    pairs_df[j,8] <- split_pairs[[i]]$Ct[1]
    pairs_df[j,9] <- split_pairs[[i]]$Ct[2]
    pairs_df[j,10] <- mean(split_pairs[[i]]$Ct)
    pairs_df[j,11] <- split_pairs[[i]]$collect_parse[1]
    j<- j+1
  }
  else {
    for (k in 2:nrow(split_pairs[[i]])){
      pairs_df[j,1] <- split_pairs[[i]]$mrn[1]
      pairs_df[j,2] <- split_pairs[[i]]$Ct[k] - split_pairs[[i]]$Ct[1]
      pairs_df[j,3] <- as.numeric(difftime(split_pairs[[i]]$collect_parse[k],split_pairs[[i]]$collect_parse[1], units="hours"))
      pairs_df[j,4] <- split_pairs[[i]]$Age[1]
      pairs_df[j,5] <- split_pairs[[i]]$Sex[1]
      pairs_df[j,6] <- split_pairs[[i]]$Method[k]
      pairs_df[j,7] <- (split_pairs[[i]]$Method[1]==split_pairs[[i]]$Method[k])
      pairs_df[j,8] <- split_pairs[[i]]$Ct[1]
      pairs_df[j,9] <- split_pairs[[i]]$Ct[k]
      pairs_df[j,10] <- (split_pairs[[i]]$Ct[1]+split_pairs[[i]]$Ct[k])/2
      pairs_df[j,11] <- split_pairs[[i]]$collect_parse[1]
      j<- j+1
    }
  }
}
colnames(pairs_df)<-c("MR", "CT_diff", "Time_diff", "Age", "Sex", "Method", "Method_match", "CT1", "CT2", "CT_mean", "first_collect_date") 

#QC on Method matching
any(!pairs_df$Method_match)

# how many in first 24 hours 
sum(length(pairs_df$CT_diff[pairs_df$Time_diff<24])) #19 in first 24 hours

sort(pairs_df$first_collect_date[pairs_df$Time_diff<48]) #7/40 are after May
hist(pairs_df$first_collect_date[pairs_df$Time_diff<48], "months", freq = TRUE)
pairs_df$first_collect_date

```

plotting the pair with time on x, CT diff on y, fill by method
```{r plot-pairs, echo=TRUE, warning=FALSE, message = FALSE}

# color scheme
dark_col <- brewer.pal(6,"Dark2")

# layering in CT differences for each machine
# Cephied 5.299954
# Hologic 4.299973
# Abbott  1.859981
ggplot(data=pairs_df, aes(x=Time_diff, y=CT_diff, col=Method))+
  annotate("rect", xmin = 0, xmax = 48, ymin = -1.859981, ymax = 1.859981,
        alpha = .3, fill=dark_col[1]) +
  annotate("rect", xmin = 0, xmax = 48, ymin = 1.859981, ymax = 4.299973,
        alpha = .3, fill=dark_col[3]) +
  annotate("rect", xmin = 0, xmax = 48, ymin = -1.859981, ymax = -4.299973,
        alpha = .3, fill=dark_col[3]) +
  annotate("rect", xmin = 0, xmax = 48, ymin = 4.299973, ymax = 5.299954,
        alpha = .3, fill=dark_col[2]) +
  annotate("rect", xmin = 0, xmax = 48, ymin = -4.299973, ymax = -5.299954,
        alpha = .3, fill=dark_col[2]) +
  geom_point()+
  theme_minimal() +
  scale_color_manual(values=dark_col)+
  xlab("Hours since first swab")+
  ylab("Difference in Ct Value")+
  xlim(c(0,48)) +
  ggtitle("Swab intrasample variability within first 48 hours,\nshaded areas corresponding to Ct Diff btwn mortality for each method")
```

4) Observe how many specimen pairs cross the Youden Index for their respective machine
```{r youden-pairs, echo=TRUE, warning=FALSE, message = FALSE}
# grabbing only those pairs within 48 hours
pairs_df_48<- pairs_df[pairs_df$Time_diff<48,]
dim(pairs_df_48)

# Abbott 
## number of pairs
sum(pairs_df_48$Method=="Abbott") #8
## number cross threshold
sum((pairs_df_48$CT1[pairs_df_48$Method=="Abbott"]<youden_coord_abbott$threshold &
       pairs_df_48$CT2[pairs_df_48$Method=="Abbott"]>youden_coord_abbott$threshold) | 
      (pairs_df_48$CT1[pairs_df_48$Method=="Abbott"]>youden_coord_abbott$threshold &
         pairs_df_48$CT2[pairs_df_48$Method=="Abbott"]<youden_coord_abbott$threshold)) #3
idx_abbott_youden <- which((pairs_df_48$CT1[pairs_df_48$Method=="Abbott"]<youden_coord_abbott$threshold &
       pairs_df_48$CT2[pairs_df_48$Method=="Abbott"]>youden_coord_abbott$threshold) | 
      (pairs_df_48$CT1[pairs_df_48$Method=="Abbott"]>youden_coord_abbott$threshold &
         pairs_df_48$CT2[pairs_df_48$Method=="Abbott"]<youden_coord_abbott$threshold)) #3

# Cepheid 
## number of pairs
sum(pairs_df_48$Method=="Cephied") #14
## number cross threshold
sum((pairs_df_48$CT1[pairs_df_48$Method=="Cephied"]<youden_coord_cephied$threshold &
       pairs_df_48$CT2[pairs_df_48$Method=="Cephied"]>youden_coord_cephied$threshold) | 
      (pairs_df_48$CT1[pairs_df_48$Method=="Cephied"]>youden_coord_cephied$threshold &
         pairs_df_48$CT2[pairs_df_48$Method=="Cephied"]<youden_coord_cephied$threshold)) #2
idx_Cephied_youden <- which((pairs_df_48$CT1[pairs_df_48$Method=="Cephied"]<youden_coord_cephied$threshold &
       pairs_df_48$CT2[pairs_df_48$Method=="Cephied"]>youden_coord_cephied$threshold) | 
      (pairs_df_48$CT1[pairs_df_48$Method=="Cephied"]>youden_coord_cephied$threshold &
         pairs_df_48$CT2[pairs_df_48$Method=="Cephied"]<youden_coord_cephied$threshold)) #3
# Hologic 
## number of pairs
sum(pairs_df_48$Method=="Hologic") #18
## number cross threshold
sum((pairs_df_48$CT1[pairs_df_48$Method=="Hologic"]<youden_coord_hologic$threshold &
       pairs_df_48$CT2[pairs_df_48$Method=="Hologic"]>youden_coord_hologic$threshold) | 
      (pairs_df_48$CT1[pairs_df_48$Method=="Hologic"]>youden_coord_hologic$threshold &
         pairs_df_48$CT2[pairs_df_48$Method=="Hologic"]<youden_coord_hologic$threshold)) #3

idx_holo_youden <- which((pairs_df_48$CT1[pairs_df_48$Method=="Hologic"]<youden_coord_hologic$threshold &
       pairs_df_48$CT2[pairs_df_48$Method=="Hologic"]>youden_coord_hologic$threshold) | 
      (pairs_df_48$CT1[pairs_df_48$Method=="Hologic"]>youden_coord_hologic$threshold &
         pairs_df_48$CT2[pairs_df_48$Method=="Hologic"]<youden_coord_hologic$threshold)) #3



pairs_df_48$CT1[pairs_df_48$Method=="Hologic"][idx_holo_youden]

## highlighting the pairs which crossed the Youden index 
pairs_df_48$`Crossed Youden` <- "No"
pairs_df_48$`Crossed Youden`[pairs_df_48$Method=="Abbott"][idx_abbott_youden] <- "Yes"
pairs_df_48$`Crossed Youden`[pairs_df_48$Method=="Cephied"][idx_Cephied_youden] <- "Yes"
pairs_df_48$`Crossed Youden`[pairs_df_48$Method=="Hologic"][idx_holo_youden] <- "Yes"

#What is the mortality status of Crossed Youden samples?
crossed_pairs <- pairs_df_48[pairs_df_48$`Crossed Youden`=="Yes",]
crossed_pairs$direction <- ifelse(crossed_pairs$CT_diff>0, "up", "down")
idx_crossed_youden <- which(mort_dat$MRN %in% pairs_df_48[pairs_df_48$`Crossed Youden`=="Yes",1])
mort_dat[idx_crossed_youden,]

# layering in CT differences for each machine
# Cephied 5.299954
# Hologic 4.299973
# Abbott  1.859981
Cross_youden_plot <- ggplot(data=pairs_df_48, aes(x=Time_diff, y=CT_diff, col=Method, size=`Crossed Youden`, shape=`Crossed Youden`))+
  annotate("rect", xmin = 0, xmax = 48, ymin = -1.859981, ymax = 1.859981,
        alpha = .3, fill=dark_col[1]) +
  annotate("rect", xmin = 0, xmax = 48, ymin = 1.859981, ymax = 4.299973,
        alpha = .3, fill=dark_col[3]) +
  annotate("rect", xmin = 0, xmax = 48, ymin = -1.859981, ymax = -4.299973,
        alpha = .3, fill=dark_col[3]) +
  annotate("rect", xmin = 0, xmax = 48, ymin = 4.299973, ymax = 5.299954,
        alpha = .3, fill=dark_col[2]) +
  annotate("rect", xmin = 0, xmax = 48, ymin = -4.299973, ymax = -5.299954,
        alpha = .3, fill=dark_col[2]) +
  geom_point()+
  theme_minimal() +
  scale_size_discrete(range = c(1.5, 3))+
  scale_color_manual(values=dark_col)+
  xlab("Hours since first swab")+
  ylab("Difference in Ct Value")+
  xlim(c(0,48)) +
  ggtitle("Swab intrasample variability within first 48 hours,\nshaded areas corresponding to Ct Diff btwn mortality for each method")
Cross_youden_plot
#ggsave(plot = Cross_youden_plot, filename = "Fig2_youden_cross.pdf", width=6.5, height=4, useDingbats=FALSE)

```
Out of the 40 pairs, 8 crossed the youden index point for predicting mortality (20%). Lending evidence to being cautionary when using Ct value to predict mortality. 

Session Information:
```{r session-Info, echo=TRUE, warning=FALSE, message = FALSE}
sessionInfo()
```